### 2013-07-31 dit-paper_ieee.tex
made the bibliography style ieeetran

%% bare_conf.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

%
\documentclass[conference]{IEEEtran}
\usepackage{cite}
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
\fi
\usepackage[cmex10]{amsmath}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{fixltx2e}
\usepackage{url}

% added
\usepackage{amssymb}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{stmaryrd}

\newtheorem{property}{Property}
\newtheorem{assumption}{Assumption}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{remark}{Remark}
\newtheorem{constraint}{Constraint}
\newtheorem{corollary}{Corollary}
\newtheorem{condition}{Condition}

\def\QED{\mbox{\rule[0pt]{1.5ex}{1.5ex}}}
\def\proof{\noindent{{\textbf{Proof}: }}}
\def\endproof{\hspace*{\fill}~\QED\par\endtrivlist\unskip \vspace{1\baselineskip}}

\newcommand{\dbf}[1]{\operatorname{dbf}(#1)}

\algsetup{
  indent=1em,
  linenosize=\small,
  linenodelimiter=\
}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
\title{The Space of Feasible Execution Times for Asynchronous Periodic Task Systems using
Definitive Idle Times}

\author{
\IEEEauthorblockN{Thomas Chapeaux and Paul Rodriguez}
\IEEEauthorblockA{Universit\'{e} Libre de Bruxelles\\
tchapeau@ulb.ac.be, paurodri@ulb.ac.be}
\and
\IEEEauthorblockN{Laurent George and ???}
\IEEEauthorblockA{ECE Paris\\
lgeorge@ece.fr}
}
\maketitle

\begin{abstract}
Sensitivity analysis for real-time systems provides for efficient methods to determine feasibility
conditions  in the case of small changes in task parameters, such as when porting a system to a
different platform. The C-space, presented in \cite{bini2004schedulability,george2009characterization}, describes the
set of execution times for which a system is feasible. This paper extends previous works to
generalize the description of the C-space for Earliest Deadline First scheduling to
systems with initial offset. The feasibility gain offered by offsets is then
expressed as the volume ratio between the C-space of an asynchronous system and its corresponding
synchronous system.
\end{abstract}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
%\IEEEpeerreviewmaketitle

\section{Introduction}

	Abstract models for real-time systems often require the determination of the
	worst-case execution time of a job (WCET), a value highly dependent on the
	platform on which the application is deployed. Feasibility conditions allow to
	decide whether a system can be scheduled correctly (i.e. without violating any
	timeliness constraint) based on task parameters, including the WCET.

	The \emph{C-space} of a real-time system is, intuitively,
	the set of WCET values for which this system would be feasible. A
	precise and efficient description of the C-space of a system would allow to quickly decide
	whether or not it is feasible on a given platform.

	In section~\ref{sct:relatedWorks}, the description of the C-space of synchronous
	constrained systems, which has been mostly covered by \cite{george2009characterization}, is revisited.
	This description is based on the DBF feasibility condition described in
	\cite{baruah1999generalized,baruah1990algorithms} and is found with exponential complexity using
	the notion of definitive idle times (DIT).

	In section~\ref{sct:asyncDIT}, we present an extension of the concept of DIT adapted to the
	asynchronous case called first periodic DIT
	($t_d$) along with an algorithm to determine its existence and value for any system,
	then we show the influence of some factors on its existence.

	In Section~\ref{sct:asyncCspace}, we show how to describe the C-space of an asynchronous
	system based on the DBF test, we show why the interval $[t_d, t_d + H]$ is
	sometimes sufficient to consider and we give a numerical example.

	Finally, in section~\ref{sct:expCspaceGain}, we use a simulation to quantify the feasibility
	gain of asynchronous system in terms of volume of the C-space.

\section{Related works}
\label{sct:relatedWorks}
	\subsection{Model}

		We consider a discrete timeline and tasks systems $\tau$ comprised of periodic concrete
		tasks with constrained deadlines on uniprocessor platforms. Each task
		$\tau_i$ is represented by a tuple $(O_i, C_i, D_i, T_i)$ where $O_i$ is the offset value,
		$C_i$ the WCET, $D_i$ the relative deadline and $T_i$ the period.
		The constrained deadline property implies that $\forall i,\; D_i \leq T_i$. In
		the following, we differentiate between \emph{synchronous systems} (where $O_i
		= 0 \; \forall i$) and \emph{asynchronous systems} (which are not synchronous).

		For those systems, the preemptive \emph{Earliest Deadline First} (EDF) algorithm, in
		which jobs are prioritized according to their absolute deadline, has been
		shown to be optimal \cite{liu1973scheduling}, which means it
		correctly schedules any feasible system within that group.

		Furthermore, the following notations are used:
		\begin{itemize}
			\item $\tau = \{\tau_1, \cdots, \tau_{n}\}$
			\item $H = \operatorname{lcm}(T_1, \cdots, T_{n})$
			\item $O_{max} = \max (O_1, \cdots, O_{n})$
			\item An \emph{idle time} is an instant $t$ at which each job that arrived
			\emph{strictly} before $t$ has finished its execution.
			\item A \emph{busy period} is a time interval between two consecutive idle times in
			which at least one job is executing.
			\item The interval between the initial time ($t=0$) and the first
			idle time is called the \emph{first busy period}.
		\end{itemize}

	\subsection{The DBF feasibility test}

	We recall the necessary and sufficient condition explained in \cite{baruah1999generalized, baruah1990algorithms}, based on the demand-bound function.

	\begin{definition}
		The \textbf{demand-bound function (DBF)}
		defined for a task set $\tau$ in a time interval $[t_1, t_2]$ and denoted $\dbf{t_1, t_2}$, is
		equal to the maximal cumulated execution time of jobs of $\tau$ contained in the
		closed interval $[t_1, t_2]$.
	\end{definition}

	Mathematically,
	\[
		\dbf{t_1, t_2} = \sum_{i=0}^{n-1} n_i(t_1, t_2) \, C_i
	\]
	where $n_i(t_1, t_2)$ is the number of jobs of task $i$ which arrival times
	and deadlines are both in the closed interval $[t_1, t_2]$.

	The values of the $n_i$ are given by \cite{baruah1999generalized}
	\[
		n_i(t_1, t_2) =
		\left\llbracket
			\left\lfloor
				\frac{t_2 - O_i - D_i}{T_i}
			\right\rfloor -
			\left\lceil
				\frac{t_1 - O_i}{T_i}
			\right\rceil + 1
		\right\rrbracket_0
	\]
	where $\llbracket x \rrbracket_0$ stands for $\max \{ 0, x \}$.

	A necessary and sufficient condition of feasibility can easily be written
	based on the DBF \cite{baruah1999generalized}

	\begin{theorem}
	\[
		\begin{array}{c}
			\{\tau_1, ..., \tau_n\} \: \text{is feasible}  \iff \\
			\dbf{t_1, t_2} \leq t_2 - t_1 \; \forall \: 0 \leq t_1 \leq t_2
		\end{array}
	\]
	\end{theorem}

	\subsection{C-space of synchronous systems}
	\label{sct:cspaceDescr}

			The authors of \cite{george2009characterization} give the following definition of the C-space (based on \cite{bini2004schedulability}):
			\begin{definition}
				The \textbf{C-space} of a task system $\tau$ is a region of $n$ dimensions (where each dimension denotes the possible $C_i$ of a task of $\tau$) such that for any vector $C = \{ C_1, \cdots, C_{n}\}$ in it, $\tau$ is feasible.
			\end{definition}

      As explained in \cite{nemhauser1988integer}, a region of $\mathbb{N}_0^n$ such as the C-space can be described by a \emph{polytope}, or a list of parametric linear constraints (called the \emph{formulation}). The constraints must be conditions which can unequivocally decide if a point is included in the region or not. In the sequel we give a formulation for the C-space in $\mathbb{N}_0^n$ (For a formulation in $\mathbb{N}^n$, we can add $C_i \geqslant 1 \; \forall i$ to the list of constraints).

      The DBF test gives us a list of inequations on the WCET values of the task set which together are a necessary and sufficient condition of feasibility. They can all be written in the following way:
      \[
        \sum_{i=0}^{n-1} n_i(t_1, t_2) \, C_i \leq t_2 - t_1 \; \forall \: 0 \leq t_1 \leq t_2
      \]
      Note that for a given $t$, we have that the $n_i(t)$ are
      constant (w.r.t. the $C_i$). Those inequations are thus indeed linear.

      However, this description contains an infinite number of constraints, most of
      which are redundant. While this does not harm the precision of the
      test, a smaller description is preferred as it can greatly
      reduce the computation time of practical feasibility tests. In Section~\ref{sct:removeRedundancy} a method to reduce the size of this set of constraints is given.

  \subsection{Removing redundant constraints}
    \label{sct:removeRedundancy}

  In the synchronous case, is has been shown in \cite{baruah1990algorithms} that the constraints corresponding
  to every interval $[0, t]$ (with $t$ a job deadline happening before $H$) are
  necessary and sufficient to test the feasibility of a system.

  Even then, the number of constraints to consider is exponential in the number
  of tasks. While it has been shown that constraints that cover intervals ending
  after the end of the first busy period are also redundant, this value depends on the $C_i$ and is
  therefore not usable for the purposes of this paper. The
  following concept is thus introduced, which is similar to the end of the busy period but does not depend on the execution times.

  \begin{definition}
    A \textbf{definitive idle time} (DIT) is a time $t$ such that every job
    released strictly before instant $t$ has its absolute deadline before or at instant $t$.
  \end{definition}

  \begin{definition}
    The earliest DIT occurring in the system strictly after $t=0$ is called the
    \textbf{first DIT} and is noted $t_d$.
  \end{definition}

  \begin{remark} (from the definitions):
    \begin{itemize}
      \item Contrary to other idle times, DITs are independent of the
      scheduling or the execution times of the jobs.
      \item In a feasible system, every DIT is an idle time.
      \item In a feasible system, the first DIT happens after the end of the
      first busy period.
      \item As any $kH$ with $k \geqslant 0$ is a DIT in synchronous systems, we have at least $t_d
      \leqslant H$ in this case.
    \end{itemize}
  \end{remark}

  The value of the first DIT can be found by solving the following modular equation system:
  \[
    t_d = \min
    \left( t \; s.t. \; \forall i:
      \left\{
        \begin{array}{c}
          t > 0 \\
          t \equiv a_i \; (mod \; T_i) \\
          a_i \in [D_i, T_i]
        \end{array}
      \right.
    \right)
  \]
  The authors of \cite{george2009characterization} give a method to solve this system for cases in which periods are pairwise coprime.

  For synchronous systems in which the execution times are not known, the first DIT $t_d$
  is our ``best guess'' at an upper limit on the end of the busy period. As it was sufficient
  to check the DBF test in intervals $[0, t]$ before the end of the busy period, it is
  slightly more pessimistic but still correct to use the following condition:
  \[
  	\begin{array}{c}
  		\dbf{0,t} \leqslant t \; \forall t \leqslant t_d \\
  		\iff \text{The system is feasible}
  	\end{array}
  \]

  To summarize, we reduced the description of the C-space from an infinite number
  of constraints to only those where $t_1 = 0$, $t_2 \leqslant t_d$ and
  values of $t_2$ correspond to job deadlines. Note that the number of
  constraints is still exponential in the number of tasks.

\section{DIT in asynchronous systems}
  \label{sct:asyncDIT}

  \subsection{Definition}

    In an asynchronous system, the first DIT may happen before $O_{max}$, i.e. before all tasks are in the system. We thus introduce the following notion:

    \begin{definition}
      The \textbf{first periodic DIT} (FPDIT) of a system is the earliest DIT occurring
      strictly after $O_{max}$.
    \end{definition}

    By a slight abuse of notation, it will also be noted $t_d$. Note that in the
    synchronous case the first periodic DIT is also the first DIT.

  \subsection{Existence condition}
  \label{sct:FPDITexist}

		As in the synchronous case, the problem of existence of the FPDIT
		can be written as a system of modular equations:
		\[
			\begin{array}{l}
				\exists ? \; t_d \; s.t. \; \forall i \; :\\
				\left\{
					\begin{array}{l}
						t_d > O_i \\
						t_d - O_i \equiv a_i \; (mod \; T_i) \\
						a_i \in [D_i, T_i]
					\end{array}
				\right.
			\end{array}
		\]

		This system is guaranteed to have at least one solution modulo $H$ for given
		values of $a_i$ if the following condition is respected:
		\[
			a_i + O_i \equiv a_j + O_j \; (\operatorname{mod} \; \operatorname{gcd}(T_i,
			T_j)) \; \forall i,j
		\]
		which means that there could be systems with no FPDIT.
		We now present an example of such a system.\\

		\begin{center}
			\begin{tabular}{|r|c|c|c|c|}
				\hline
							& $O_i$ & $C_i$ & $D_i$ & $T_i$ \\ \hline
				$\tau_1$ 	& 0 	& $C_1$ & 3 	& 4\\ \hline
				$\tau_2$ 	& 2 	& $C_2$ & 7 	& 8\\ \hline
			\end{tabular}
		\end{center}

		Here, the possible values of $a_1$ are 3 and 4, and the possible values
		of $a_2$ are 7 and 8. For a FPDIT to exist, one of the following
		conditions must be fulfilled:
		\[
			\left[
			\begin{array}{c}
				3 + 0 \equiv 7 + 2 \; (\operatorname{mod} \; 4 ) \\
				3 + 0 \equiv 8 + 2 \; (\operatorname{mod} \; 4 ) \\
				4 + 0 \equiv 7 + 2 \; (\operatorname{mod} \; 4 ) \\
				4 + 0 \equiv 8 + 2 \; (\operatorname{mod} \; 4 ) \\
			\end{array}
			\right.
		\]
		However neither is, which means the system has
		no FPDIT (and no other DIT than $t=0$).

	\subsection{Solving the modular system}
		The existence conditions of	Section~\ref{sct:FPDITexist} can directly be used to calculate the value of
		the FPDIT (if any), because for a given set of $a_i$ values they are
		very similar to an instance of the Chinese Remainder Problem (CRP), which is
		easily solved using Gauss's CRP solving method for non-coprime moduli
		\cite{gauss1965disquisitiones}.

		However this is very inefficient as there are exponentially many different
		combinations of $a_i$ values in the number of tasks, and resolving
		exponentially many CRP is prohibitively slow. A better method for finding the
		FPDIT consists in splitting Gauss's algorithm into two separate parts, as
		shown in Algorithm~\ref{alg:multiACRP} (note that $x^{-1} \; (mod \; y)$
		means the modular inverse of $x$ in the naturals modulo $y$).

		\begin{algorithm}
			\caption{Modified Gauss Algorithm}
			\label{alg:multiACRP}
			\begin{algorithmic}[1]
				\REQUIRE $A$ is a list of the sets of possible $a_i$ per task
				\REQUIRE $H$ is the hyperperiod of the system

				\STATE $sumchunks \leftarrow \text{empty list}$
				\FORALL{$\tau_j \in \tau$}
					\STATE $taskTerms \leftarrow \text{empty list}$
					\STATE $Mj \leftarrow \frac{H}{T_j}$
					\STATE $invMj \leftarrow Mj^{-1} (mod \; T_j)$
					\FORALL{$a_i \in A(\tau_j)$}
						\STATE $\text{append} \; (a_i \cdot Mj \cdot invMj) \; \text{to} \; taskTerms$
					\ENDFOR
					\STATE $\text{append} \; taskTerms \; \text{to} \; sumchunks$
				\ENDFOR

				\STATE $results \leftarrow \text{empty list}$
				\STATE append $0$ to $results$
				\FORALL{$\tau_j \in \tau$}
					\STATE $newresults \leftarrow \text{empty list}$
					\FORALL{$a_i \in A(\tau_j)$}
						\FORALL{$r \in results$}
							\STATE $\text{append} \; (r + sumchunks(\tau_j,i)) \; \text{to} \; newresults$
						\ENDFOR
					\ENDFOR
					\STATE $results \leftarrow newresults$
				\ENDFOR
				\RETURN $r \; (mod \; H)$ for all $r \in results$
			\end{algorithmic}
		\end{algorithm}

		The global structure of the original algorithm is simply a sum of factors, each term being dependent on which $a_i$ value is chosen. A major
		speedup is achieved by first computing all possible terms for each task
		then creating all possible DIT values by summing the terms in all possible
		ways. The number of results is still exponential in the number of tasks, but
		the only operation that is repeated exponentially many times is now a sum.
		This allows us to profit from a high chance of having DITs when the deadlines
		are close to the WCETs (and far from the periods), but with only a slight
		increase in computation time as the sizes of the $[D_i, T_i]$ intervals grow.

	\subsection{Simulation}
	\label{sct:noFPDITsimu}

	This simulation aims to quantify how much system have a FPDIT and which factor can influence this quantity.

	We thus checked the existence condition on 10000 systems generated by:
	\begin{itemize}
		\item Generating utilization values with the \texttt{UUniFast} algorithm described in \cite{bini2005measuring}, with parameters
		\begin{itemize}
			\item $U_{tot}$ (system utilization) chosen uniformly between 0.25 and 0.75
			\item $n$ (number of tasks) equal to 3, 4 and 5.
		\end{itemize}
		\item Choosing the $T_i$ uniformly between 5 and 20, which also gives us the $C_i = \lfloor U_i \, T_i \rfloor$
		\item Choosing the $O_i$ with a normal distribution of mean $T_{min}$ and standard deviation $\frac{T_{max} - T_{min}}{2}$
		\item Choosing the $D_i$ uniformly in the interval $[T_i - CDF(T_i - C_i), T_i]$, where $CDF$ varies between 0 and 1 and is called the \emph{Constrained Deadline Factor}. It allows us to generate systems where the deadlines are closer to the period.
	\end{itemize}

	The number of tasks and the constrained Deadline Factor is shown to have a direct influence on the density of FPDIT. Indeed, Fig~\ref{fig:noFPDIT} show the percentage of systems having no FPDIT for different values of CDF.

	\begin{figure}[h]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{python-simulation/plots/nofpdit_3.png}
	\end{center}
	\caption{Evolution of the existence of FPDIT with the CDF}
	\label{fig:noFPDIT}
	\end{figure}

	We observe that the closer we are to the implicit deadline case ($CDF =
	0$), the fewer systems have a FPDIT. This is easily explained by the fact that
	a lower CDF means more idle instants for each task, and thus more opportunities
	for DITs to happen. Similarly, as all tasks must be synchronized for a DIT to
	happen, a higher number of tasks results in a decrease in FPDIT density.

\section{C-space of asynchronous systems}
\label{sct:asyncCspace}

	\subsection{Description using DBF}

		The description we presented in Section \ref{sct:cspaceDescr} was general, and
		is thus still relevant. Therefore the C-space is still accurately described by
		the following infinite number of constraints:

		\[
			\sum_{i=0}^{n-1} n_i(t_1, t_2)
			\, C_i \leq t_2 - t_1 \; \forall \: 0 \leq t_1 \leq t_2
		\]

		However, the method used in the synchronous case to remove redundancy does not
		apply here. A more general approach is presented in the following section.

	\subsection{Removing redundant constraints}

		The first two following sections contain well-known results of feasibility analysis.
		The third one is a new result based on the FPDIT and the last section contains a general
		but heavy method for removing constraints redundancies.

		\subsubsection{Job arrival and deadline}
			Consider a time interval $[t_1, t_2]$. If $t_2$ is an instant without any
			job deadline, then by definition the value of $\dbf{t_1, t_2}$ will be equal to
			$\dbf{t_1, t_2^*}$, where $t_2^*$ is the latest deadline before $t_2$.

			A similar argument can be made if $t_1$ is an instant without any job arrival
			(and $t_1^*$ is the earliest arrival after $t_1$). We can thus restrict
			ourselves to intervals where $t_1$ is an instant with at least one job
			arrival, and $t_2$ is an instant with at least one job deadline.

		\subsubsection{Feasibility interval}
			First let us recall how an asynchronous system behaves under an optimal
			scheduler, for example EDF.

			\begin{figure}[h]
				\[
					\begin{array}{r||c|c|c|l}
									& \text{Incomplete} & \text{Transient}	& \text{Stationary} & \cdots \\
						\text{Name} & \text{period} 	& \text{period} 	& \text{period}  	& \\ \hline
						\text{size} & O_{max} 			& H 				& H 				& \cdots \\ \hline
						t \geqslant & 0 				& O_{max} 			& O_{max} + H 		& \cdots
					\end{array}
				\]
				\begin{center}
				\caption{Behavior of an asynchronous system under EDF.}
				\label{fig:asyncBehavior}
				\end{center}
			\end{figure}

			As shown in Fig. \ref{fig:asyncBehavior}, the system begins in the
			\emph{incomplete period}, where all tasks are not in the system. Then, after
			$O_{max}$ time units, every tasks is in the system and the \emph{transient
			period} begins. Then, after $O_{max} + H$ time units, the system is faced
			with the same pattern of arrival as in $O_{max}$. However, because in
			$O_{max} + H$ every task is in the system and in $O_{max}$ some were missing,
			the state of the system will be different. The \emph{stationary period} then
			begins. At instant $O_{max} + 2H$ it is guaranteed that the system is in the
			exact same situation it was at instant $O_{max} + H$.

			For this reason, as shown by \cite{leung1982complexity}, it is sufficient (but pessimistic) to
			check for
			intervals included in the interval $[O_{max}, O_{max} + 2H]$, which has an exponential length
			in the number of tasks. A better interval is given in \cite{choquet2004minimal} using the
			notion of cyclic idle times (denoted $t_c$), which are dependent of the execution
			time and therefore cannot be used directly. The interval is $[0, t_c + H]$ and is valid for
			every $t_c$.

			In the following section, we extend the latter interval using the previously defined notion
			of periodic DIT, a particular case of cyclic idle times which does not depend on the execution
			times.

\subsubsection{Using the first periodic DIT}

The FPDIT, if it exists, must occur in the transient period. Indeed, it cannot occur in the incomplete period by definition ($t_d > O_{max}$) and if it occurred in the $k^{th}$ stationary period, another DIT should have occurred in the transient period at instant $t_d - k H$ (a contradiction).

As the FPDIT happens when all tasks are in the system, we know that the system will arrive in the exact same state one hyperperiod later ($t_d + H$) (A rigorous proof is given in \cite{choquet2004minimal} for any cyclic idle time). We can thus restrict the DBF test to intervals included in the interval $[t_d, t_d + H]$ if the FPDIT exists. This will always be a better interval than $[O_{max}, O_{max} + 2H]$ and if the first periodic DIT does not exist (which can be verified as explained in Section \ref{sct:FPDITexist}), the latter interval can be used.

\subsubsection{Redundancy as an integer linear problem}

The general problem of redundancy of a new linear constraint $C \, X \leq d$ w.r.t. a set of previous linear constraints $A \, X \leq B$ in an ILP can itself be written as the following ILP:

\begin{figure}[h]
$$z = \max C \, X$$
s.t.
\[
\begin{array}{rccc}
  A & X &\leq & B \\
  C & X &\leq & d + 1
\end{array}
\]
with
\[
  \begin{array}{ccc}
    A & : & [n,k] \text{ matrix}\\
    B & : & [n,1] \text{ matrix}\\
    C & : & [n,1] \text{ matrix}\\
    X & : & [1,n] \text{ matrix}\\
    d & : & \text{integer}
  \end{array}
\]
\caption{Redundancy of $CX \leq d$ w.r.t. $A X \leq B$ as an ILP}
\end{figure}

Indeed, if the maximization returns $z=d+1$, it means that there are some integer values accessible with the previous constraints that the new constraint forbids. On the other hands, if the maximization returns $z \leq d$, it means that the limitation expressed in the new constraint is already enforced by the previous constraints. Therefore $z \leq d$ is a necessary and sufficient condition of redundancy of the new constraint.

Such a linear problem can be solved by e.g. the simplex algorithm and a branch-and-cut approach. However, in our case the problem is not exactly the same as we have a set of constraints and we want to find all redundant constraints within. A naive method would be to test each constraint against all the others. We present a similar method with an efficient preprocessing when most constraints are redundant in Algorithm \ref{alg:prunRedun}.

\begin{algorithm}
\caption{Removing redundancy from CSPACE}
\label{alg:prunRedun}
  \begin{algorithmic}[1]
    \STATE $CSPACE$ : set of constraints
    \STATE $S \leftarrow \emptyset$
    \STATE \COMMENT{$1^{st}$ pass: test each $cstr$ against previous ones}
    \FOR{$cstr$ \textbf{in} $CSPACE$}
      \IF {$cstr$ is not redundant w.r.t. $S$}
        \STATE $S \leftarrow cstr$
      \ENDIF
    \ENDFOR
    \STATE \COMMENT{$2^{nd}$ pass: test each $cstr$ against every other}
    \FOR{$cstr$ \textbf{in} $S$}
      \STATE Remove $cstr$ from $S$
      \IF {$cstr$ is not redundant w.r.t. $S$}
        \STATE Put $cstr$ back into $S$
      \ENDIF
    \ENDFOR
    \RETURN{S}
  \end{algorithmic}
\end{algorithm}

After doing this, no redundant constraints are left. However, this algorithm
takes an exponential time in the number of constraints. An efficient initial
description of $CSPACE$, as given by the methods described in previous
sections, is therefore preferred.

To summarize, we first reduced the number of constraints from every possible
interval $[t_1, t_2]$ to the intervals where $t_1$ is an arrival time and
$t_2$ is a deadline in $[t_d, t_d + H]$ if $t_d$ exists, or in
$[O_{max}, O_{max} + 2H]$ otherwise. This set of constraints is a
correct description of the C-Space, but it is not efficient. Redundant
contraints are removed through a linear integer approach.

\section{Numerical Example}

Consider the following task system

    \begin{center}
    \begin{tabular}{|r|c|c|c|c|}
     \hline
      & $O_i$ & $C_i$ & $D_i$ & $T_i$ \\
     \hline
     $\tau_1$ & 8 & $C_1$ & 7 & 15\\
     \hline
     $\tau_2$ & 0 & $C_2$ & 2 & 5\\
     \hline
    \end{tabular}
    \end{center}

The FPDIT happens at $t=15$ and we have $H = 15$. The study interval using the FPDIT is thus $[15, 30]$, which contains 11 intervals $[a,d]$ with $a$ being the arrival of a job, $d$ being the deadline of a job, and $15 \leqslant a < d \leqslant 30$. Those intervals are
$[15, 17]$, $[15, 22]$, $[15, 27]$, $[15, 30]$, $[20, 22]$, $[20, 27]$, $[20, 30]$, $[23, 27]$, $[23, 30]$, $[25, 27]$ and $[25, 30]$. For comparison, the original study interval of $[O_{max}, O_{max} + 2H]$ is $[8, 38]$ which yields 57 intervals.

For each $[a, d]$, we have a constraint $$\sum_{i=0}^{n-1} n_i(a, d) \, C_i \leqslant d - a$$ which give a description of the C-space.

We now apply Algorithm \ref{alg:prunRedun} to remove redundancies, leaving us with the following constraints:
$$
\left\{
  \begin{array}{ccccc}
    & & C_2 & \leqslant & 2 \\
    C_1 & + & C_2 & \leqslant & 7
  \end{array}
\right.
$$
The C-space thus contains 11 points $(C_1, C_2) = (1, 1)$, $(2, 1)$, $(3, 1)$, $(4, 1)$, $(5, 1)$, $(6, 1)$, $(1, 2)$, $(2, 2)$, $(3, 2)$, $(4, 2)$ and $(5, 2)$.

Let us now consider the same system but with a synchronous pattern of arrival.

    \begin{center}
    \begin{tabular}{|r|c|c|c|c|}
     \hline
      & $O_i$ & $C_i$ & $D_i$ & $T_i$ \\
     \hline
     $\tau_1$ & 0 & $C_1$ & 7 & 15\\
     \hline
     $\tau_2$ & 0 & $C_2$ & 2 & 5\\
     \hline
    \end{tabular}
    \end{center}

The first DIT happens at time $t_d = 7$. The intervals to consider are thus of the form $[0, t]$ where $t$ is a deadline happening before or at time $t_d$. This yields two intervals ($[0, 2]$ and $[0, 7]$) and the following constraints describing the C-space, none of them being redundant:
$$
\left\{
  \begin{array}{ccccc}
    & & C_2 & \leqslant & 2 \\
    C_1 & + & 2 C_2 & \leqslant & 7
  \end{array}
\right.
$$
This time, the C-space only contains 8 points: $(1, 1)$, $(2, 1)$, $(3, 1)$, $(4, 1)$, $(5, 1)$, $(1, 2)$, $(2, 2)$, $(3, 2)$.

Fig. \ref{fig:cspaceComp} shows the two C-spaces in the plane. The dark gray region is the C-space of the synchronous system and the light gray region is the surface gain given by the asynchronous system.

\begin{figure}[h]
\begin{center}
  \includegraphics[width=0.5\textwidth]{figs/cspace_example.png}
  \caption{C-space of an asynchronous system and its corresponding synchronous system}
  \label{fig:cspaceComp}
\end{center}
\end{figure}

\section{Experimental C-space size gain of asynchronous systems}
\label{sct:expCspaceGain}
	One of the advantages of asynchronous systems over synchronous systems in the
	context of periodic tasks is that the critical instants of asynchronous systems
	are at most as constraining for the C-space as a synchronous instant. For this
	reason, modifying task offsets when possible during
	the design of a real-time system is a useful trick to fit a heavy periodic task set
	into a given platform.

	In Fig.~\ref{fig:sizeRatio} a number of feasible asynchronous task sets of
	three tasks without synchronous instants were generated. The size of their
	C-space (number of valid WCET vectors) was compared with that of the same
	task sets with all offsets set to zero. The systems were generated as in section~\ref{sct:noFPDITsimu}
	and the integer linear problems used to remove redundancies were solved via the GLPK tool\footnote{\url{http://www.gnu.org/software/glpk/}}.

	In other words, this figure
	evaluates the feasibility gain of adding random offsets to a synchronous task
	set. It is assumed that a more advanced strategy to assign these offsets
	would yield a greater difference.

	\begin{figure}[h]
		\begin{center}
			\includegraphics[width=0.5\textwidth]{figs/sizeratio.png}
			\caption{C-space size ratio between asynchronous and equivalent synchronous
			systems}
			\label{fig:sizeRatio}
		\end{center}
	\end{figure}

	The ratio between the size of the synchronous and asynchronous C-spaces of
	these systems roughly range from 0.95 in the worst case (high utilization, high
	CDF) to 0.85 in the best case (low utilization, low CDF). As calculating those
	sizes is rather long, the number of systems used to obtain average
	results was rather low (1000), which is the likely cause of the spikes in the
	data. Note that the value of the CDF has a greater influence on this ratio than
	the system's total utilization, but less so when the CDF is lower. It is also
	important to note that even though a 0.95 ratio may not seem much, the
	synchronous C-spaces of the systems contain many dominated WCET vectors, and
	only those that are in the border (i.e. non-dominated) actually matter to the
	designer of the system. Using the asynchronous C-space instead will relax some
	constraints and therefore add new acceptable solutions that will dominate solutions that were previously non-dominated,
	increasing the quality of the C-space border.

\section{Conclusion}

In this paper, we presented a general approach to define the C-space of a constrained periodic task set for EDF on uniform platform. Possible future works includes an extension to the study of systems with preemption costs or running on multiprocessor platforms, as the value of the DIT does not depend on preemptions or the number of processors.

Future works could also study the shape of the C-space and how it is affected by some factors. An
interesting path would be the effect of the CDF on the number of constraints describing the C-space,
as we know that in the implicit case this description is simple ($U_{tot} \leqslant 1$). While we
quantified the feasibility gain by looking at the ratio between sizes of C-space, another approach
would be to look at the ratio between the minimal distance of the border of the C-space, which would
give a lower bound of the gain given by the offsets.

% conference papers do not normally have an appendix

% use section* for acknowledgement
% \section*{Acknowledgment}

% The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
% \nocite{*}
\bibliographystyle{IEEEtran}
\bibliography{dit-paper}

% that's all folks
\end{document}